---
title: "Data science coursera track final project"
author: "Joris Van den Bossche"
date: "26 september 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = normalizePath('../'))
```

# Global options
global options and the libraries used are found in the Global folder:
```{r}
source("00_Global/libraries.R")
source("00_Global/settings.R")
```


# Reading in data
Functions are defined in this folder for downloading, sampling and loading the data into R.
```{r}
source("01_Load/load.R")
```

## Usage
Downloading the data:
```{r}
if(!file.exists("RawData/final")){
  downloadFiles()
}
```

To load a sample of the files into R, this function was created:
```{r readsample}
readTextsSample(lines = 2, lang = "en_US")
twitterTexts
```

Creating a subfolder "sample" in the "RawData" folder. This may take a few minutes.
```{r createSample, eval = FALSE}
createSampleDataDir(usePercentageOfData, seed, lang = "en_US")
```

Reading the sample data into a corpus is done using this function:
```{r createCorpus, cache = TRUE, results = 'hide'}
if(!exists("corpus") || !exists("tdm"))
  createCorpus(lang = "en_US")
```

# Exploratory Analysis
